---
title: "Glossy Buckthorn CA"
author: "Tim Szewczyk"
date: "`r Sys.Date()`"
output: github_document
---

This is a wrapper for running the glossy buckthorn cellular automata model. It sets up the environment, allows the assignment of all necessary parameters, runs the simulation, and visualizes the results.

The land cover within each cell is compositional rather than the more common hard classification. As a result, all land cover-specific parameters are computed proportionally to the land cover in a cell (i.e., lc.prop %*% K, where lc.prop is a matrix with columns for land cover and a row for each cell, and K is a vector of carrying capacities with one per land cover type).

```{r setup, include=FALSE}
if(!require(sevcheck)) devtools::install_github("Sz-Tim/sevcheck")
p_load("tidyverse", "magrittr", "stringr", "parallel", 
       "pbapply", "fastmatch", "gganimate")
theme_set(theme_bw())
source("code/ca_fn.R")
grnt <- read_csv("data/grnt_all.csv") %>%
  mutate(x=as.numeric(as.factor(.$left)),
         y=as.numeric(factor(.$top, levels=rev(levels(factor(.$top))))),
         x_y=paste(x, y, sep="_"))
lc.rct <- as.tibble(expand.grid(x=1:max(grnt$x),
                                y=1:max(grnt$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(OpI=grnt$OpI[match(.$x_y, grnt$x_y)], 
         Oth=grnt$Oth[match(.$x_y, grnt$x_y)], 
         Dec=grnt$Dec[match(.$x_y, grnt$x_y)], 
         Evg=grnt$Evg[match(.$x_y, grnt$x_y)], 
         WP=grnt$WP[match(.$x_y, grnt$x_y)], 
         Mxd=grnt$Mxd[match(.$x_y, grnt$x_y)],
         inbd=!is.na(match(.$x_y, grnt$x_y)))
lc.rct[is.na(lc.rct)] <- 0
```


```{r setPars}
##---
## set parameters
##---

  set.seed(225)
  n.sim <- 1
  g.p <- list(
    # general
    tmax=20,  # num time steps per simulation
    simple=FALSE,  # simple lambda growth vs. fruits/seeds
    dem.st=FALSE,  # include stochasticity in demography?
    sdd.st=TRUE,  # include stochasticity in SDD?
    n_cores=4,  # parallelize sdd.pr calculation
    
    # landscape
    lc.r=30,  # num rows in landscape
    lc.c=30,  # num columns in landscape
    n.lc=6,  # num land cover categories
    N.p.t0=50,  # num cells with buckthorn at t=1
    
    # demography
    K=c(750, 10, 100, 100, 300, 100),  # K for adults
    pr.s=c(0.9, 0.1, 0.6, 0.6, 0.6, 0.6),  # pre-adult survival
    pr.f=c(0.9, 0.1, 0.29, 0.23, 0.2, 0.3),  # pr(fruit)
    fec=c(200, 100, 40, 20, 20, 10),  # mean(fruit per adult)
    age.f=3,#rep(c(3,5), c(2,4)),  # age @ first fruiting (length==n.lc | 1)
    bank=TRUE,  # include seedbank?
    pr.sb=0.3,  # pr(ann.surv seed bank)
    pr.est=c(0.07, 0.01, 0.08, 0.02, 0.02, 0.03),  # pr(seedling est)
    lambda=c(1.5, 0.5, 1.0, 0.9, 1.2, 1.1),  # pop growth rate
    
    # dispersal
    sdd.max=5,#20,  # max dispersal distance (unit: cells)
    sdd.rate=1,#0.1,  # 1/mn for dispersal kernel
    pr.eat=c(0.3, 0.1, 0.2, 0.2, 0.2, 0.1),  # pr(birds eat fruit)
    bird.hab=c(.35, .35, 0.05, 0.1, 0.1, 0.05),  # bird habitat prefs
    pr.s.bird=0.6,  # seed viability post-digestion
    n.ldd=1   # num long distance dispersal events per year
  )
  
  control.p <- list(
    nTrt_grd=0.05,  # % of cells treated in each time step
    nTrt_man=0.05,  # % of cells treated in each time step
    grd.trt=c(Lit=0.005, Cov=0.01, Com=0.00001),
    man.trt=c(M=0.1, C=0.3, MC=0.8),
    t.trt=50
  )

  

##--
## initialize landscape
##--
  
  # land cover
  lc.df <- lc.rct %>% filter(y >= (max(lc.rct$y) - g.p$lc.r) & x <= g.p$lc.c) %>%
    mutate(id=row_number())
  ncell <- nrow(lc.df)
    
  
  # populations
  N.init <- pop_init(ncell, g.p, lc.df)
  
  # dispersal probabilities
  system.time(sdd.pr <- sdd_set_probs(lc.df, g.p))
```

```{r runSim}
system.time(
for(s in 1:n.sim) {
  out <- run_sim(g.p, lc.df, sdd.pr, N.init, control.p)
}
)
```


```{r plots, fig.height=8, fig.width=10}
out.sb <- log(out$N.sb+1)
out.ad <- out$N[,,max(g.p$age.f)]
N.out <- cbind(lc.df, out.ad) %>% as.tibble %>%
  gather(year, N.adult, (1:ncol(out.ad)) + ncol(lc.df)) %>%
  mutate(N.sb=c(out.sb),
         N.less5=c(out.sb > 0))
N.out$N.less5 <- N.out$N.less5 + c(out.ad > 5)
N.out$year <- str_pad(N.out$year, 3, "left", "0")

# final time step
adult.ab.fin <- ggplot(filter(N.out, year==str_pad(g.p$tmax+1, 3, "left", "0")),
                       aes(x=x, y=y, fill=N.adult)) +
  geom_tile() + scale_fill_gradient(low="white", high="red") +
  ggtitle(paste("Adult abundance. Year", g.p$tmax+1))
sb.ab.fin <- ggplot(filter(N.out, year==g.p$tmax+1), 
                    aes(x=x, y=y, fill=N.sb)) +
  geom_tile() + scale_fill_gradient(low="white", high="red") +
  ggtitle(paste("log seed abundance. Year", g.p$tmax+1))
adult.pa.fin <- ggplot(filter(N.out, year==g.p$tmax+1), 
                       aes(x=x, y=y, fill=N.adult>0)) +
  geom_tile() + scale_fill_manual(values=c("white","red")) + 
  ggtitle(paste("Adult presence. Year", g.p$tmax+1))
sb.pa.fin <- ggplot(filter(N.out, year==g.p$tmax+1), 
                    aes(x=x, y=y, fill=N.sb>0)) +
  geom_tile() + scale_fill_manual(values=c("FALSE"="white","TRUE"="red")) +
  ggtitle(paste("Seed presence. Year", g.p$tmax+1))

# gifs
gifwd <- paste0("out/gifs/", ncell, "_t", g.p$tmax, "/")
age.i <- ifelse(length(g.p$age.f)==1, 
                paste0("age", g.p$age.f, ""),
                paste0("ages", min(g.p$age.f), "-", max(g.p$age.f), ""))
adult.ab <- ggplot(N.out, aes(x=x, y=y, fill=N.adult, frame=year)) + 
  geom_tile() + scale_fill_gradient(low="white", high="red") + 
  ggtitle("Adult abundance. Year")
sb.ab <- ggplot(N.out, aes(x=x, y=y, fill=N.sb, frame=year)) + 
  geom_tile() + scale_fill_gradient(low="white", high="red") + 
  ggtitle("log seed abundance. Year")
adult.pa <- ggplot(N.out, aes(x=x, y=y, fill=N.adult>0, frame=year)) + 
  geom_tile() + scale_fill_manual(values=c("white","red")) + 
  ggtitle("Adult presence. Year")
sb.pa <- ggplot(N.out, aes(x=x, y=y, fill=N.sb > 0, frame=year)) + 
  geom_tile() + scale_fill_manual(values=c("white","red")) + 
  ggtitle("Seed presence. Year")
adult.lo <- ggplot(N.out, aes(x=x, y=y, fill=N.less5, frame=year)) +
  geom_tile() + ggtitle("Blue: Adult density ≤ 5. Year") + 
  scale_fill_gradient2(low="white", mid="blue", high="pink", midpoint=1)

# save plots
ggsave(paste0(gifwd, "Final_Ab_", age.i, "TRT_allAge_persistent.jpg"), adult.ab.fin,
       width=8, height=6, units="in")
ggsave(paste0(gifwd, "Final_SB_Ab_", age.i, "TRT_allAge_persistent.jpg"), sb.ab.fin,
       width=8, height=6, units="in")
ggsave(paste0(gifwd, "Final_PA_", age.i, "TRT_allAge_persistent.jpg"), adult.pa.fin,
       width=8, height=6, units="in")
ggsave(paste0(gifwd, "Final_SB_PA_", age.i, "TRT_allAge_persistent.jpg"), sb.pa.fin,
       width=8, height=6, units="in")
gganimate(adult.ab, paste0(gifwd, "Ab_", age.i, "TRT_allAge_persistent.gif"), 
          interval=0.2, ani.width=800, ani.height=600)
gganimate(sb.ab, paste0(gifwd, "SB_Ab_", age.i, "TRT_allAge_persistent.gif"),
          interval=0.2, ani.width=800, ani.height=600)
gganimate(adult.pa, paste0(gifwd, "PA_", age.i, "TRT_allAge_persistent.gif"),
          interval=0.2, ani.width=800, ani.height=600)
gganimate(sb.pa, paste0(gifwd, "SB_PA_", age.i, "TRT_allAge_persistent.gif"),
          interval=0.2, ani.width=800, ani.height=600)
gganimate(adult.lo, paste0(gifwd, "LoDens_", age.i, "TRT_allAge_persistent.gif"),
          interval=0.2, ani.width=800, ani.height=600)

# 
# 
# 
# N.out$lam <- as.matrix(lc.df[,4:9]) %*% g.p$lambda
# ggplot(N.out, aes(x=x, y=y, fill=lam)) + geom_tile() +
#   scale_fill_gradient2(midpoint=1)
# 
# # land cover
# p.dev <- ggplot(N.out, aes(x=x, y=y, fill=Dev)) + geom_tile() + 
#   scale_fill_gradient(low="white", high="red", limits=c(0,1))
# ggsave(paste0("out/plots/n_", ncell, "_LC_Dev.tif"), p.dev,
#          width=10, height=7)
# 
# p.oth <- ggplot(N.out, aes(x=x, y=y, fill=Oth)) + geom_tile() + 
#   scale_fill_gradient(low="white", high="gray30", limits=c(0,1))
# ggsave(paste0("out/plots/n_", ncell, "_LC_Oth.jpg"), p.oth,
#          width=10, height=7)
# 
# p.hwd <- ggplot(N.out, aes(x=x, y=y, fill=Hwd)) + geom_tile() + 
#   scale_fill_gradient(low="white", high="green3", limits=c(0,1))
# ggsave(paste0("out/plots/n_", ncell, "_LC_Hwd.jpg"), p.hwd,
#          width=10, height=7)
# 
# p.wp <- ggplot(N.out, aes(x=x, y=y, fill=WP)) + geom_tile() + 
#   scale_fill_gradient(low="white", high="orchid", limits=c(0,1))
# ggsave(paste0("out/plots/n_", ncell, "_LC_WP.jpg"), p.wp,
#          width=10, height=7)
# 
# p.evg <- ggplot(N.out, aes(x=x, y=y, fill=Evg)) + geom_tile() + 
#   scale_fill_gradient(low="white", high="darkgreen", limits=c(0,1))
# ggsave(paste0("out/plots/n_", ncell, "_LC_Evg.jpg"), p.evg,
#          width=10, height=7)
# 
# p.mxd <- ggplot(N.out, aes(x=x, y=y, fill=Mxd)) + geom_tile() + 
#   scale_fill_gradient(low="white", high="yellowgreen", limits=c(0,1))
# ggsave(paste0("out/plots/n_", ncell, "_LC_Mxd.jpg"), p.mxd,
#          width=10, height=7)
# 
# 
# # buckthorn vs land cover
# ggplot(N.out, aes(x=Dev)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Oth)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Hwd)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=WP)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Evg)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
# ggplot(N.out, aes(x=Mxd)) + 
#   geom_point(aes_string(y=paste0("t_", g.p$tmax)), alpha=0.1)
```




